---
title: "Basic Information Extraction from FARS Data"
author: "Natasa Formanova"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Basic Information Extraction from FARS Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The functions bundled in the package `farshelp` provide tools for some basic manipulation of data from the US National Highway Traffic Safety Administration's [Fatality Analysis Reporting System](https://www.nhtsa.gov/Data/Fatality-Analysis-Reporting-System-(FARS)), which is a nationwide census providing the American public yearly data regarding fatal injuries suffered in motor vehicle traffic crashes.

Example data for years 2013-2015 can be downloaded here:  
[`fars.data.zip`](https://d3c33hcgiwev3.cloudfront.net/_e1adac2a5f05192dc8780f3944feec13_fars_data.zip?Expires=1537833600&Signature=VQQlF2cnmDt5opDQkL8peyB5LOJJJr8Jjdmg2yIQQRtMYwxf-CZ9iTuD2yQXuRHjivAXJrBQAujn7ZA5VJRn--O95So8D0X5eBmxGfuPucnYJI3vZ2JZSl-TZk4I-KLCvm3aZ6wz06TJlMZhdU8VZUmTCfiUk8sG56Iq-UFv1iY_&Key-Pair-Id=APKAJLTNE6QMUY6HBC5A)

After unzipping the files nad placing them in the working directory, incidence of accidents for specified years can be extracted and summarized by year and month:

```{r, }
library(farshelp)

my_years <- fars_summarize_years(c(2013, 2014))

head(my_years)
```

Incidence of accidents for a single year in a given state can also be visualized on a map, by using the function `fars_map_state()` (note that states are identified by a numerical code):

```{r, warning = FALSE}
fars_map_state(state.num = 17, year = 2014)
```

